---
title: "Statistically Significant Richardson, Pt 1"
subtitle: "Text Mining and Literary Inference in Samuel Richardson's Clarissa"
author: "Johanna Kopecky"
date: "24 January 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      ratio: "16:9"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_light(
  base_color = "#F47B6A",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono")
)
```


```{r, echo=F, fig.width=7, fig.height=4, fig.align='center', message=FALSE, warning=FALSE}
#load necessary packages
library(ggformula)
library(tidyverse)
library(knitr)
library(tidyverse)
library(tidytext)
library(rsample)
```

<style>

  .col2 {

    columns: 2 200px;         /* number of columns and width in pixels*/

    -webkit-columns: 2 200px; /* chrome, safari */

    -moz-columns: 2 200px;    /* firefox */

  }

  .col3 {

    columns: 3 100px;

    -webkit-columns: 3 100px;

    -moz-columns: 3 100px;

  }

</style>



## Overview


* Preparing the Data
* Data from Clarissa

    -Word Frequency

    -Two-Word Phrases

* Future Work

---

## Preparing the Data


* Download full text from Project Gutenberg using gutenbergr

* Clean the data 

    -Introductions and disclaimers
    
    -"Stop words"

```{r, echo=FALSE}
get_stopwords()
```

---

## Word Frequency of Clarissa

Top 10 words that appear in Clarissa (stop words removed)

```{r, include=FALSE}
library(gutenbergr)
full_text3 <- gutenberg_download(9881)
#Calling it full_text3 because it is from Volume 3

cleaned_text3_0 <- full_text3[-c(1:297),]

cleaned_text3_1 <- cleaned_text3_0

library(tm)

tidy_book <- cleaned_text3_1 %>%
  mutate(line = row_number()) %>%
  unnest_tokens(word, text)

tidy_book %>%
  count(word, sort = TRUE)

get_stopwords()
get_stopwords(source = "smart")
```

```{r, ,echo=FALSE, fig.height=6, fig.width=10}
tidy_book %>%
  anti_join(get_stopwords(source = "smart")) %>%
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(fct_reorder(word, n), n)) +
  geom_col() +
  coord_flip()
```

---

## Two-Word Phrases

Top 10 two-word phrases that appear in Clarissa (stop words removed)

```{r, include=FALSE}
tidy_ngram <- cleaned_text3_1 %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2)
tidy_ngram

tidy_ngram %>%
  count(bigram, sort = TRUE)

tidy_ngram %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word) %>%
  count(word1, word2, sort = TRUE)

tidy_bigram <- tidy_ngram %>%
  #anti_join(get_stopwords(source = "smart")) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word) %>%
  count(word1, word2, sort = TRUE) %>%
  top_n(10)
```

```{r, echo=FALSE, fig.height=6, fig.width=10}
tidy_bigram$two_word <- paste(tidy_bigram$word1, " ", tidy_bigram$word2)

tidy_bigram[1:6,] %>% 
  ggplot(aes(fct_reorder(two_word, n), n)) +
  geom_col() +
  coord_flip() +
  xlab("phrase") +
  ylab("occurances")

```

---

## Future Work

* Compare Clarissa with a different work (Paradise Lost)
* Compare Clarissa with a similar work (Pamela)

---

## References/Acknowledgements

* Julia Silge, code and Text Mining in R

    -Silge, J., & Robinson, D. (2017). Text mining with R: A tidy approach. Sebastopol, CA: OReilly Media.

    -Silge, J. (n.d.). TEXT CLASSIFICATION WITH TIDY DATA PRINCIPLES. Retrieved from https://juliasilge.com/blog/tidy-text-classification/

* Professor Andrew Sage, advisor
* Professor Celia Barnes, advisor

---

## Questions?

* Email: johanna.r.kopecky@lawrence.edu
* Github: ShakespeareAndStats
