---
title: "Data-Scientific Programming: Assignment 9"
author: "Johanna Kopecky"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```


```{r}
library(tidyverse)
library(rvest)
library(robotstxt)
```

##### This assignment pertains to creating webscraping. Refer to notesheets 9.1 and 9.2 for help and examples as you work through the assignemnt. 

In this assignment, we'll scrape data on smart watches from [ebay](https://www.ebay.com/b/Smart-Watches/178893/bn_152365?LH_BIN=1&rt=nc). 

##### 1) Use the `paths_allowed()` function to check that you can scrape data from ebay. Use the url "http://www.ebay.com". 

```{r}
paths_allowed("https://www.harpconnection.com/harpstore/harp-ConcertPedal.html")
```

##### 2) Read in the url. 

```{r}
page1 <- read_html("https://www.harpconnection.com/harpstore/harp-ConcertPedal.html")
```

##### 3) Scrape in all of the following:

a) name of the harp

```{r}
name <- page1 %>%
  html_nodes("#stacks_in_2_page39 .THCCenter a") %>%
  html_text() %>%
  as.character()

head(name)
```

b) price

```{r}
price <- page1 %>%
  html_nodes(".THCsmall") %>%
  html_text() %>%
  str_remove_all("New from ") %>%
  str_remove_all("\\$") %>%
  str_remove_all("\\*") %>%
  str_remove_all("\\,") %>%
  as.numeric()

head(price)
```

c) company that made the harp 

```{r}
company <- page1 %>%
  html_nodes(".plusplus .THCCenter") %>%
  html_text() %>%
  str_split("by ") %>%
  map(2, .default=NA) %>%
  as.character()

head(company)
```

d) number of strings

```{r}
strings <- page1 %>%
  html_nodes(".THCHarpModel tr:nth-child(2) .paragraph") %>%
  html_text() %>%
  str_split("\\-") %>%
  map(1, .default=NA) %>%
  str_remove_all("\n") %>%
  trimws() %>%
  as.numeric()

head(strings)
```

e) type of harp

```{r}
type <- page1 %>%
  html_nodes(".THCHarpModel tr:nth-child(2) .paragraph") %>%
  html_text() %>%
  str_split("String") %>%
  map(2, .default=NA) %>%
  str_remove_all("\n") %>%
  trimws() %>%
  as.character()

head(type)
```



####################################################################################


##### 2) Read in the url. 

```{r}
page2 <- read_html("https://www.harpconnection.com/harpstore/harp-SemiPetitePedal.html")
```

##### 3) Scrape in all of the following:

a) name of the harp

```{r}
name2 <- page2 %>%
  html_nodes("#stacks_in_2_page39 .THCCenter a") %>%
  html_text() %>%
  as.character()

head(name2)
```

b) price

```{r}
price2 <- page2 %>%
  html_nodes(".THCsmall") %>%
  html_text() %>%
  str_remove_all("New from ") %>%
  str_remove_all("\\$") %>%
  str_remove_all("\\*") %>%
  str_remove_all("\\,") %>%
  as.numeric()

head(price2)
```

c) company that made the harp 

```{r}
company2 <- page2 %>%
  html_nodes(".plusplus .THCCenter") %>%
  html_text() %>%
  str_split("by ") %>%
  map(2, .default=NA) %>%
  as.character()

head(company2)
```

d) number of strings

```{r}
strings2 <- page2 %>%
  html_nodes(".THCHarpModel tr:nth-child(2) .paragraph") %>%
  html_text() %>%
  str_split("\\-") %>%
  map(1, .default=NA) %>%
  str_remove_all("\n") %>%
  trimws() %>%
  as.numeric()

head(strings2)
```

e) type of harp

```{r}
type2 <- page2 %>%
  html_nodes(".THCHarpModel tr:nth-child(2) .paragraph") %>%
  html_text() %>%
  str_split("String") %>%
  map(2, .default=NA) %>%
  str_remove_all("\n") %>%
  trimws() %>%
  as.character()

head(type2)
```



#####################################################################################

```{r}
page3 <- read_html("https://www.harpconnection.com/harpstore/harp-ClassicLever.html")
```



##### 3) Scrape in all of the following:

a) name of the harp

```{r}
name3 <- page3 %>%
  html_nodes(".plusplus a") %>%
  html_text() %>%
  as.character()

head(name3)
```

b) price

```{r}
price3 <- page3 %>%
  html_nodes(".THCsmall") %>%
  html_text() %>%
  str_split("Slightly")  %>%
  map(1, .default=NA) %>%
  str_remove_all("New from ") %>%
  str_remove_all("\\$") %>%
  str_remove_all("\\*") %>%
  str_remove_all("\\\"") %>%
  str_remove_all("\\,") %>%
  as.numeric()

head(price3)
```

c) company that made the harp 

```{r}
company3 <- page3 %>%
  html_nodes(".plusplus .THCCenter") %>%
  html_text() %>%
  str_split("by ") %>%
  map(2, .default=NA) %>%
  as.character()

head(company3)
```

d) number of strings

```{r}
strings3 <- page3 %>%
  html_nodes(".THCHarpModel tr:nth-child(2) .paragraph") %>%
  html_text() %>%
  str_split("\\-") %>%
  map(1, .default=NA) %>%
  str_remove_all("\n") %>%
  trimws() %>%
  as.numeric()

head(strings3)
```

e) type of harp

```{r}
type3 <- page3 %>%
  html_nodes(".THCHarpModel tr:nth-child(2) .paragraph") %>%
  html_text() %>%
  str_split("String") %>%
  map(2, .default=NA) %>%
  str_remove_all("\n") %>%
  str_remove_all("with Standard Pedal Harp") %>%
  str_remove_all("\\(") %>%
  trimws() %>%
  as.character()

head(type3)
```

####################################################################################

```{r}
page4 <- read_html("https://www.harpconnection.com/harpstore/harp-CelticHarps.html")
```

##### 3) Scrape in all of the following:

a) name of the harp

```{r}
name4 <- page4 %>%
  html_nodes(".plusplus a") %>%
  html_text() %>%
  as.character()

head(name4)
```

b) price

```{r}
price4 <- page4 %>%
  html_nodes(".THCsmall") %>%
  html_text() %>%
  str_split("Slightly")  %>%
  map(1, .default=NA) %>%
  str_remove_all("New from ") %>%
  str_remove_all("\\$") %>%
  str_remove_all("\\*") %>%
  str_remove_all("\\\"") %>%
  str_remove_all("\\,") %>%
  as.numeric()

head(price4)
```

c) company that made the harp 

```{r}
company4 <- page4 %>%
  html_nodes(".plusplus .THCCenter") %>%
  html_text() %>%
  str_split("by ") %>%
  map(2, .default=NA) %>%
  as.character()

head(company4)
```

d) number of strings

```{r}
strings4 <- page4 %>%
  html_nodes(".THCHarpModel tr:nth-child(2) .paragraph") %>%
  html_text() %>%
  str_split("\\-") %>%
  map(1, .default=NA) %>%
  str_remove_all("\n") %>%
  trimws() %>%
  as.numeric()

head(strings4)
```

e) type of harp

```{r}
type4 <- page4 %>%
  html_nodes(".THCHarpModel tr:nth-child(2) .paragraph") %>%
  html_text() %>%
  str_split("String") %>%
  map(2, .default=NA) %>%
  str_remove_all("\n") %>%
  trimws() %>%
  as.character()

head(type4)
```

#########################################################################

```{r}
harps1 <- tibble(name, price, strings, company, type)

head(harps1)
```

```{r}
harps2 <- tibble(name2, price2, strings2, company2, type2)

head(harps2)
```

```{r}
harps3 <- tibble(name3, price3, strings3, company3, type3)

head(harps3)
```

```{r}
harps4 <- tibble(name4, price4, strings4, company4, type4)

head(harps4)
```

```{r}
harps02 <- full_join(x=harps3, y=harps4, by=c("name3"="name4"))

head(harps02)
```






